Iconic representations of molluscs through the ages

### 2012-10-29 large_number_squeezing_with_atoms.tex
Finish the appendix

This system can be solved analytically, and expressions for quadrature, absolute and relative number squeezing can be derived. We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}.
If we define
\begin{eqnarray}
s &=& \sin(\Omega (t_3 - t_2)) \\
c &=& \cos(\Omega (t_3 - t_2)) \\
\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} / \hbar \\
A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
D &=& AB^* - A^*B
\end{eqnarray}
then the expectation values for number, number variance, and number difference variance after appling the final coupling field for a time $\tau = t_3-t_2$ are given by
\begin{eqnarray}
N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D \\
%
N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D \\
%
{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2) \nonumber \\
       && + i c s D (1-2 n_a c^2 -2 n_b s^2)  \nonumber\\
       && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ) \nonumber\\
%
{\mathrm{Var}} [ N_b](t_3) &=&  n_a s^2 + n_b c^2 + 2 n_a n_b c^2 s^2 \nonumber \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2 ) \nonumber \\
       && + i c s D (-1+2 n_a s^2 + 2 n_b c^2)  \nonumber \\
       && + 2 i c^3 s n_b (e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - e^{i(\lambda_{bb} - \lambda_{ab})} A B^* ) \nonumber \\
       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ) \\
%
{\mathrm{Var}} [ N_a - N_b](t_3) &=& n_a + n_b + 4 i c s D (n_a - n_b)(s^2 - c^2) \nonumber \\
       && + 4 c^2 s^2 (2 n_a n_b +D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} A_2^* B_2 - e^{-i(\lambda_{aa} - \lambda_{bb})} A_2 B_2^* ) \nonumber \\
       && + 4 i c s (c^2 - s^2) (n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B 
                     + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^*)
\end{eqnarray}
In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, e.g. ${\mathrm{Var}}[N_a]=N_a$, an no number squeezing is possible.

@@

We are now in a position to evaluate the number variance of the system throughout period of the final coupling. We first consider the number variance of the $\hat{a}$ field, which is defined as
{\mathrm{Var}}[N_a] = \langle \hat{a}^{\dagger}_3 \hat{a}_3 \hat{a}^{\dagger}_3 \hat{a}_3 \rangle - \langle \hat{a}^{\dagger}_3 \hat{a}_3 \rangle ^2.
\label{eqNavariance}

@@

Clearly, to calculate (\ref{eqNavariance}) we will need the expectation values of terms quartic and quadratic in $\hat{a}_2$ and ${\hat{b}_2}$. Specifically, we require $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \rangle$, $\langle \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$,  $\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \rangle$, $\langle \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$ , $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 \rangle$, and $\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$, as well as their Hermitian conjugates.

@@

Next we tackle $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$. Using (\ref{eqa2}) and (\ref{eqb2}) and the fact that $\hat{a}$ and $\hat{b}$ commute, we have
\begin{eqnarray}
\hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 &=& \hat{a}^{\dagger}_1 
  e^{i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
  \hat{b}_1 \hat{a}^{\dagger_1}
  e^{i (\lambda_{ab}-\lambda_{bb}) \hat{b}^{\dagger}_1 \hat{b}_1}
  e^{i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
  \hat{b}_1 \\
%
&=& \hat{a}^{\dagger}_1 \hat{a}^{\dagger}_1 
    e^{2i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
    e^{2i (\lambda_{ab}-\lambda_{bb}) \hat{b}^{\dagger}_1 \hat{b}_1}
    \hat{b}_1 \hat{b}_1 e^{i(\lambda_{aa}-\lambda_{bb})}
\end{eqnarray}
where we have made multiple uses of the identity 
\begin{equation}
\hat{a} e^{i x \hat{a}^{\dagger} \hat{a}} = e^{i x \hat{a}^{\dagger} \hat{a}} \hat{a} e^{ix}
\end{equation}
which follows from (\ref{eqefeidentity}). Finally, applying (\ref{eqExpectationValueOfExponential}) we obtain
\begin{equation}
\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle = \alpha^{*2} \beta^2 e^{i(\lambda_{aa}-\lambda_{bb})} \exp[|\alpha|^2(e^{2i (\lambda_{aa} - \lambda_{bb})} -1)] \exp[|\beta|^2(e^{-2i (\lambda_{bb} - \lambda_{bb})} -1)] 
\end{equation}

then the expectation values for number and number squared at time $t=t_3$ are given by
\begin{eqnarray}
\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \rangle &=& n_a c^2 + n_b s^2 + i c s (A B^* - A^* B) \\
%
\langle \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \rangle &=& n_a s^2 + n_b c^2 - i c s (A B^* - A^* B) \\
%
\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle &=& (n_a^2 + n_a) c^4 + (n_b^2 + n_b) s^4 \nonumber \\
&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2] \nonumber \\
&& + i c^3 s [2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})}] A B^* - 2 n_a e^{i(\lambda_{aa}-\lambda_{ab})} A^* B + A B^* - A^* B] \nonumber \\
&& + i c s^3 [2 n_b e^{i(\lambda_{bb}-\lambda_{ab})}] A B^* - 2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})} A^* B + A B^* - A^* B] \\
%
\langle \hat{b}^{\dagger}(t_3)  \hat{b}(t_3) \hat{b}^{\dagger}(t_3)  \hat{b}(t_3) \rangle &=& (n_b^2 + n_b) c^4 + (n_a^2 + n_a) s^4 \nonumber \\
&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2 - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^*] \nonumber \\
&& + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* - A^* B + A B^*] \nonumber \\
&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* - A^* B + A B^*] 
\end{eqnarray}

\begin{eqnarray}
\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle &-& \langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \rangle^2 \nonumber \\
%
&=& \langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle \\
&-& n_a^2 c^4 + n_b^2 s^4 - c^2 s^2 (A B^* - A^* B)^2 + 2 i n_a c^3 s (A B^* - A^* B) + 2 i n_b c s^3 (A B^* - A^* B) + n_a n_b c^2 s^2 \\
%
\langle \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \rangle &=& n_a s^2 + n_b c^2 - i c s (A B^* - A^* B) \\
%
\langle \hat{a}^{\dagger}(t_3)  \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}(t_3) \rangle &=& (n_a^2 + n_a) c^4 + (n_b^2 + n_b) s^4 \nonumber \\
&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2] \nonumber \\
&& + i c^3 s [2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})}] A B^* - 2 n_a e^{i(\lambda_{aa}-\lambda_{ab})} A^* B + A B^* - A^* B] \nonumber \\
&& + i c s^3 [2 n_b e^{i(\lambda_{bb}-\lambda_{ab})}] A B^* - 2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})} A^* B + A B^* - A^* B] \\
%
\langle \hat{b}^{\dagger}(t_3)  \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \hat{b}(t_3) \rangle &=& (n_b^2 + n_b) c^4 + (n_a^2 + n_a) s^4 \nonumber \\
&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2 - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^*] \nonumber \\
&& + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* + A^* B - A B^*] \nonumber \\
&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* + A^* B - A B^*] 
\end{eqnarray}

>>>>

This system can be solved analytically, and expressions for quadrature, absolute and relative number squeezing can be derived. We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}.
If we define
\begin{eqnarray}
s &=& \sin(\Omega (t_3 - t_2)) \label{eqsDef} \\
c &=& \cos(\Omega (t_3 - t_2)) \\
\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} / \hbar \\
A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
D &=& AB^* - A^*B \label{eqDDef}
\end{eqnarray}
then the expectation values for number, number variance, and number difference variance after appling the final coupling field for a time $\tau = t_3-t_2$ are given by
\begin{eqnarray}
N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D \\
%
N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D \\
%
{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2) \nonumber \\
       && + i c s D (1-2 n_a c^2 -2 n_b s^2)  \nonumber\\
       && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ) \nonumber\\
%
{\mathrm{Var}} [ N_b](t_3) &=&  n_a s^2 + n_b c^2 + 2 n_a n_b c^2 s^2 \nonumber \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2 ) \nonumber \\
       && + i c s D (-1+2 n_a s^2 + 2 n_b c^2)  \nonumber \\
       && + 2 i c^3 s n_b (e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - e^{i(\lambda_{bb} - \lambda_{ab})} A B^* ) \nonumber \\
       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ) \\
%
{\mathrm{Var}} [ N_a - N_b](t_3) &=& n_a + n_b + 4 i c s D (n_a - n_b)(s^2 - c^2) \nonumber \\
       && + 4 c^2 s^2 (2 n_a n_b +D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} A_2^* B_2 - e^{-i(\lambda_{aa} - \lambda_{bb})} A_2 B_2^* ) \nonumber \\
       && + 4 i c s (c^2 - s^2) (n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B 
                     + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^*) \label{eqNumDiffVariance}
\end{eqnarray}
In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, e.g. ${\mathrm{Var}}[N_a]=N_a$, an no number squeezing is possible.

@@

We are now in a position to evaluate the number variance of the system throughout the period of the final coupling. The number variances of the fields are defined as
{\mathrm{Var}}[N_a] &=& \langle \hat{a}^{\dagger}_3 \hat{a}_3 \hat{a}^{\dagger}_3 \hat{a}_3 \rangle - \langle \hat{a}^{\dagger}_3 \hat{a}_3 \rangle ^2 \label{eqNavariance} \\
{\mathrm{Var}}[N_b] &=& \langle \hat{b}^{\dagger}_3 \hat{b}_3 \hat{b}^{\dagger}_3 \hat{b}_3 \rangle - \langle \hat{b}^{\dagger}_3 \hat{b}_3 \rangle ^2
\label{eqNbvariance}

@@

\hat{b}^{\dagger}_3 \hat{b}_3 &=& \hat{b}^{\dagger}_2 \hat{b}_2 c^2 +  \hat{a}^{\dagger}_2 \hat{a}_2 s^2 + i c s e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 - i c s e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 \\
\hat{b}^{\dagger}_3 \hat{b}_3 \hat{b}^{\dagger}_3 \hat{b}_3 &=& \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 c^4 + \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 s^4 \nonumber \\
&& + c^2 s^2 [ \hat{b}^{\dagger}_2 \hat{b}_2 + \hat{a}^{\dagger}_2 \hat{a}_2 + 4 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{a}^{\dagger}_2 \hat{a}_2 - e^{2 i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 -e^{2 i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 ] \nonumber \\
&& + i c^3 s [2 e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 - 2 e^{-i \phi} \hat{b}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{a}_2 + e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 - e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 ] \nonumber \\
&& + i c s^3 [2 e^{i \phi} \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{a}_2 - 2 e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}_2 + e^{i \phi} \hat{a}^{\dagger}_2 \hat{b}_2 - e^{-i \phi} \hat{b}^{\dagger}_2 \hat{a}_2 ] 
Clearly, to calculate (\ref{eqNavariance}) and (\ref{eqNbvariance}) we will need the expectation values of terms quartic and quadratic in $\hat{a}_2$ and ${\hat{b}_2}$. Specifically, we require $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \rangle$, $\langle \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$,  $\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \rangle$, $\langle \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$ , $\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$, $\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 \rangle$, and $\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$, as well as their Hermitian conjugates.

@@

Next we tackle $\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle$. Using (\ref{eqa2}) and (\ref{eqb2}) and the fact that $\hat{a}$ and $\hat{b}$ commute, we have
\hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 &=& \hat{a}^{\dagger}_1 
  e^{i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
  \hat{b}_1 \hat{a}^{\dagger_1}
  e^{i (\lambda_{ab}-\lambda_{bb}) \hat{b}^{\dagger}_1 \hat{b}_1}
  e^{i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
  \hat{b}_1 \\
&=& \hat{a}^{\dagger}_1 \hat{a}^{\dagger}_1 
    e^{2i (\lambda_{aa}-\lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1}
    e^{2i (\lambda_{ab}-\lambda_{bb}) \hat{b}^{\dagger}_1 \hat{b}_1}
    \hat{b}_1 \hat{b}_1 e^{i(\lambda_{aa}-\lambda_{bb})}
where we have made multiple uses of the identity 
\hat{a} e^{i x \hat{a}^{\dagger} \hat{a}} = e^{i x \hat{a}^{\dagger} \hat{a}} \hat{a} e^{ix}
which follows from (\ref{eqefeidentity}). Finally, applying (\ref{eqExpectationValueOfExponential}) we obtain
\langle \hat{a}^{\dagger}_2 \hat{a}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle = \alpha^{*2} \beta^2 e^{i(\lambda_{aa}-\lambda_{bb})} \exp[|\alpha|^2(e^{2i (\lambda_{aa} - \lambda_{bb})} -1)] \exp[|\beta|^2(e^{-2i (\lambda_{bb} - \lambda_{bb})} -1)] 
Using the same techniques one can show that
\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{a}_2 \hat{a}_2 \rangle &=& |\alpha|^2 \alpha \beta^* e^{-i(\lambda_{aa}-\lambda_{ab})} \exp[|\alpha|^2(e^{-i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2(e^{i (\lambda_{bb} - \lambda_{ab})} -1)] \\
\langle \hat{a}^{\dagger}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}_2 \rangle &=& |\beta|^2 \alpha^{*} \beta e^{-i(\lambda_{bb}-\lambda_{ab})} \exp[|\alpha|^2(e^{i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2(e^{-i (\lambda_{bb} - \lambda_{ab})} -1)].
Employing the definitions (\ref{eqsDef})--(\ref{eqDDef}) we obtain
\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \rangle &=& n_a c^2 + n_b s^2 + i c s (A B^* - A^* B) \\
\langle \hat{b}^{\dagger}(t_3) \hat{b}(t_3) \rangle &=& n_a s^2 + n_b c^2 - i c s (A B^* - A^* B) \\
\langle \hat{a}^{\dagger}(t_3) \hat{a}(t_3) \hat{a}^{\dagger}(t_3)  \hat{a}(t_3) \rangle &=& (n_a^2 + n_a) c^4 + (n_b^2 + n_b) s^4 \nonumber \\
&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2] \nonumber \\
&& + i c^3 s [2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})}] A B^* - 2 n_a e^{i(\lambda_{aa}-\lambda_{ab})} A^* B + A B^* - A^* B] \nonumber \\
&& + i c s^3 [2 n_b e^{i(\lambda_{bb}-\lambda_{ab})}] A B^* - 2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})} A^* B + A B^* - A^* B] \\
\langle \hat{b}^{\dagger}(t_3)  \hat{b}(t_3) \hat{b}^{\dagger}(t_3)  \hat{b}(t_3) \rangle &=& (n_b^2 + n_b) c^4 + (n_a^2 + n_a) s^4 \nonumber \\
&& + c^2 s^2 [n_a +n_b +4 n_a n_b - e^{-i(\lambda_{aa}-\lambda_{bb})} B_2^* A_2 - e^{i(\lambda_{aa}-\lambda_{bb})} B_2 A_2^*] \nonumber \\
&& + i c^3 s [2 n_b e^{-i(\lambda_{bb}-\lambda_{ab})}] A^* B - 2 n_b e^{i(\lambda_{bb}-\lambda_{ab})} A B^* - A^* B + A B^*] \nonumber \\
&& + i c s^3 [2 n_a e^{i(\lambda_{aa}-\lambda_{ab})}] A^* B - 2 n_a e^{-i(\lambda_{aa}-\lambda_{ab})} A B^* - A^* B + A B^*] 
When substituted into (\ref{eqNbvariance}) and (\ref{eqNbvariance}) these expecation values give the absolute number variances stated in Section \ref{secTwoModeAnalytic}. To calculate the number difference variance, we note
{\mathrm{Var}}[N_a-N_b] = {\mathrm{Var}}[N_a] + {\mathrm{Var}}[N_b] - 2\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2\rangle + 2\langle \hat{a}^{\dagger}_2  \hat{a}_2 \rangle \langle \hat{b}^{\dagger}_2 \hat{b}_2\rangle \label{eqNumDiffVarianceAppendix}.
Substituting the expectation values calculated earlier in this Appendix into (\ref{eqNumDiffVarianceAppendix}) and simplifying yields the expression (\ref{eqNumDiffVariance}) given in Section \ref{secTwoModeAnalytic}.
